Modeling trend in temperature volatility using generalized LASSO

### 2018-05-15 nips-2018/paper.tex
updates on paper

%Arash
\usepackage{amsmath,amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\DeclareMathOperator*{\argmin}{argmin}
\usepackage{siunitx}

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}

@@

@@

@@

Inspired by the $\ell_1$-trend filtering algorithm, we propose a
non-parametric model for estimating the variance of a time-series. To
this end, we assume that at each time step $t$, there is a hidden
variable $h_t$ such that conditioned on $h_t$ the observations $y_t$
are independent normal variables with zero mean and variance
$\exp(h_t)$. The negative log-likelihood of the observed data in this
model is $l(y\given h) \propto -\sum_{t=1}^Th_t - y_t^2e^{-h_t}$. Crucially,
we assume that the hidden variables $h_t$ vary smoothly. To impose
this assumption, we estimate $h_t$ by solving the penalized, negative
log-likelihood: 
\begin{equation}
\min_h -l(y\given h)+\lambda \norm{ Dh }_1
\label{eq:l1tf_var}
\end{equation}
 where $D$ has the same structure as above.

@@

As with~\eqref{eq:l1tf}, one can solve~\eqref{eq:l1tf_var} using the
PDIP algorithm (as in, e.g.,
\texttt{cvxopt}~\citep{andersen_cvxopt:_2013}). First, we note that
this is a generalized LASSO problem
\citep{tibshirani_solution_2011}. The dual of a generalized LASSO with
the objective $f(x)+\lambda \norm{ Dx }_1$ is:  
\begin{align}
\min_\nu&\quad f^*(-D^\top\nu) & \mbox{s.t.}&\quad \norm{ \nu }_\infty \le \lambda
\end{align}
 where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x
 u^\top x-f(x)$. It is simple to show that 
\begin{equation}
f^*(u)=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
\label{eq:conj}
\end{equation}
Writing
\begin{equation}
  r_w(v,\mu_1,\mu_2):=
  \begin{bmatrix}
    \nabla f^*(-D^\top v) + D(v-\lambda \one)^\top \mu_1 -
    D(v+\lambda \one)^\top \mu_2\\
    -\mu_1(v-\lambda\one)+\mu_2(v + \lambda\one) -w^{-1}\one
  \end{bmatrix}
  =
  \begin{bmatrix}
  0\\0\end{bmatrix}
\label{eq:resid}
\end{equation}
for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the
$\ell_\infty$ constraint, as $w\rightarrow\infty$,
the solution to this nonlinear system reduces to the KKT
conditions. Therefore, the PDIP method takes Newton steps to solve the
system for a series of increasing values of $w$. \autoref{alg:pdip}
provides the details. 
\begin{algorithm}[tb]
  \caption{PDIP for $\ell_1$ variance estimation}
  \label{alg:pdip}
  \begin{algorithmic}
    \REQUIRE $\lambda>0$, $w>0$, $\nu\leftarrow 0$, $\mu_1\leftarrow
    0$, $\mu_2\leftarrow 0$, $J\in\mathbb{Z}^+$, $\{w_k\}$
    \COMMENT{Initialization} 
    \FOR[Central path]{$k=1,2,\ldots$}
    \FOR[Newton updates]{$j=1,\ldots,J$}
    \STATE Solve $A [\Delta\nu\; \Delta\mu_1\; \Delta\mu_2]^\top=r_{w_k}$ to find the search direction.
    \STATE $\quad\quad A$ is the Jacobian of $r_w$ in \eqref{eq:resid}.
    \STATE Update $[\nu^{j+1}\; \mu^{j+1}_1\; \mu^{j+1}_2] \leftarrow
    [\nu^{j}\; \mu^{j}_1\; \mu^{j}_2] + [\Delta\nu\; \Delta\mu_1\;
    \Delta\mu_2]$. 
    \ENDFOR
   \ENDFOR
   \RETURN $h=\log\frac{y^2}{1+D^\top\nu}$ \attn{Is this right?
     Explicit form of $A$}
  \end{algorithmic}
\end{algorithm}

@@

The first term in the objective is proportional to the negative
log-likelihood, the second is the temporal penalty for the
time-series at each location $(i,j)$, while the third and fourth,
penalize the difference between the estimated variance of two
vertically and horizontally adjacent points, respectively. The spatial
component of this
penalty is a special case of trend
filtering on graphs~\citep{wang_trend_2016} which penalizes the difference
between the estimated values of the signal on the connected
nodes. As before, we can write
\eqref{eq:l1tf_var_st} in matrix form where $h$ is an $T\times
S$ vector and $D$ is replaced by $D_{TS}$. Then, as we
have two different tuning parameters for the temporal and spatial
components, we write $\lambda =\left[\lambda_1\one_T^\top,\;
  \lambda_2\one_{S}^\top\right]^\top$ and $\Lambda =\texttt{diag}(\lambda)$ leading to 
% let $h$ be a
% vector whose first $T$ entries are $h_{11t}$ for $t=1,...,T$, its next
% $T$ entries are $h_{21t}$ and so on, then the optimization problem in
% the matrix form is as
\begin{equation}
\min_h -l(y\given h)+ \norm{\Lambda D_{TS}h }_1.
\label{eq:l1tf_var_st_mat}
\end{equation}
% where $\Lambda = \mathbf{I}_{Tn_rn_c}\left[\lambda_1\one_T^\top,\;
%   \lambda_2\one_{n_rn_c}^\top\right]$ and 
% $D_{total}^t=[D_{t}^t|D_{s}^t]$, where $n_t$ and $n_s$ are the number
% of rows of $D_{t}$ and $D_{s}$, respectively (see below). The matrix
% $D_t$ is the following block-diagonal matrix and corresponds to the
% temporal penalty: 
% \begin{equation}
% D_t=\begin{bmatrix}
% D &  & \\ 
%  & \ddots & \\
%  &  & D
% \end{bmatrix}
% \label{eq:d_t_matrix}
% \end{equation}
%  where $D$ was defined in \eqref{eq:d_matrix}. The number of the
%  diagonal blocks is equal to the grid size $n_r \times n_c$. Each row
%  of the matrix $D_s$ corresponds to one spatial constraint in
%  \eqref{eq:l1tf_var_st}. For example, the first $T$ rows correspond to
%  $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to
%  $|h_{11t}-h_{12t}|$, and so on. 
% \attn{Detailed discussion of the PDIP to here}
The dual of this problem is 
$\min_\nu f^*(-D_{ST}^\top\nu)$ subject to 
$| \nu_k | \le \Lambda_{kk},\; k=1,\ldots,ST$,
 where $f^*$ is given in \eqref{eq:conj}.

@@

For a spatial grid of size $S$ and $T$ time steps, $D_{ST}$ will have
$3n_rn_c+Tn_c+2Tn_r$ rows and $ST$ columns. For a $1^\circ\times
1^\circ$ grid over the
entire northern hemisphere and daily data over 10 years, we have
$n_r=90$, $n_c=180$, $T=3650$ and so $D_{ST}$ has approximately
$10^8$ columns and $10^8$
rows. In each
step of \autoref{alg:pdip}, we need to solve a linear system of
equations in $A$ which depends on $D_{ST}^\top D_{ST}$ (see \citep{boyd_convex_2004}
equation 11.54). Therefore, 
applying the PDIP directly is infeasible for our data.\footnote{We note that this
  is a highly structured and sparse matrix, but, unlike trend
  filtering alone, it is not banded. We are unaware of general
  linear algebra techniques for inverting such matrix, despite our
  best efforts.}

@@

General (unconstrained) optimization problems $\min_z
\sum_i f_i(z)$, $z\in\mathbb{R}^n$ can be rewritten equivalently using a
collection of local variables, $x_i\in \mathbb{R}^{n_i}$, and a
constraint. Following the notation of 
\citep{boyd_distributed_2011}, let $k=\mathscr{G}(i,j)$ which
means that the $j^\text{th}$ entry of $x_i$ is $z_k$ (or
$(x_i)_j=z_k$) and define $\tilde{z}_i \in \mathbb{R}^{n_i}$ by
$(\tilde{z}_i)_j=(x_i)_j$. Then the original unconstrained
optimization problem is equivalent to the following constrained
problem
$\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)$ subject to $\tilde{z}_i=x_i$.
Now, we can apply ADMM to the augmented Lagrangian of this
problem: 
\begin{align}
x_i&\leftarrow \argmin_{x_i} f_i(x_i) + (u_i)^\top x_i +
     (\rho/2) \norm{ x_i-\tilde{z}_i }_2^2\\
\label{eq:ADMM_steps}
z_k&\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j\\
u_i&\leftarrow u_i + \rho (x_i-\tilde{z}_i)
\end{align}
Here, $S_k$ is the number of local variable entries that correspond to
$z_k$, and $u_i$ are the Lagrange multipliers.  The $x$-update is
often a simple modification as the original optimization, though over only a small
partition, the $z$-update simply averages over collections of
analogous local variables while the $u$ update enforces sharing. Both
the $x$ and $u$ updates are parallelizable over the partition of local
variables.
\attn{I don't much like this explanation or notation. Seems unclear}

To solve the optimization problem \eqref{eq:l1tf_var_st_mat} or
\eqref{eq:l1tf_var_st} using this method, we need to two address two
questions: first, how to choose the 
local variables $x_i$, and second, how to solve the optimization
problem for updating these variables (the first line of
\eqref{eq:ADMM_steps}).

In \autoref{fig:data_cube}, the global variable $z$ is represented
as a cube (using the subset of the US as an example). We decompose $z$
into sub-cubes as shown in the figure by white lines. It is easy to see
that with this definition of $x_i$, the objective
\eqref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where
$f_i(x_i)=-l(y_i\given x_i)+\norm{\Lambda_{(i)} D_{(i)}x_i}$, and
$\Lambda_{(i)}$ and $D_{(i)}$ contain the temporal and spatial
penalties corresponding to $x_i$ only.
\attn{I would like to be able to write all this in terms of $h$
  instead of $x$}

\begin{figure}[tb]
  \centering
  \includegraphics[height=.2\textheight]{Figures/data_cube}
    \caption{The cube represents the global variable $z$ in space and
      time. The sub-cubes specified by the white lines are
      $x_i$. \attn{Can we use $h_i$ here instead of $x_i$?}}
    \label{fig:data_cube}
\end{figure}

By this definition of $x_i$, the update step for $x_i$ is the
following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) +
(u_i^m)^t x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.

We solve this using the PDIP method. To this end we first need to
compute the dual of the problem. It can be shown that the dual of this
problem is: 
%TODO: write the appendix
\begin{align}
\min_\nu &f_i^*(-D_i^t\nu_i) & s.t. &| (\nu_i)_j | \le (\Lambda_i)_j
\label{eq:dual_x_update}
\end{align}
 where:
\begin{align}
f^*_i(q)&=\sum_j (q_i)_j (w_i)_j -(w_i)_j-(y_i)_j^2 e^{-(w_i)_j} - 
(\rho/2)((w_i)_j-(\alpha_i^m)_j)^2\\
\alpha_i^m &=\hat{z}_i^m-u_i^m \label{eq:conj_x_update}\\
(w_i)_j&=\mathscr{W}\bigg(\frac{(y_i)_j^2}{\rho}\exp
         \bigg[\frac{1-q_j-\rho(\alpha_i^m)_j}{\rho}\bigg]\bigg)- 
\frac{1-q_j-\rho(\alpha_i^m)_j}{\rho}
\end{align}
\attn{I'm not quite sure what's going on here. These need to match the
  generic discussion in \eqref{eq:ADMM_steps}. Or are these just
  defining some other variables for convenience? This needs to be
  clarified and incorporated into \autoref{alg:ADMM}.}

In these equations, $\mathscr{W}$ is the \textit{Lambert function}
\citep{corless_lambertw_1996}. To use the PDIP method, we also need to
compute the gradient and Hessian of $f_i^*(-D_i^t\nu_i)$. This
involves computing the derivatives of the Lambert functions. The
primal solution $x_i$ can be obtained from the dual solution $\nu_i$
in \eqref{eq:dual_x_update} by setting $x_i=w_i$ where $w_i$ is
defined in the last equation of \eqref{eq:conj_x_update}.

@@

\attn{Need a good dummy variable here, it can't be $u$.} Consider the generic optimization problem
$\min_x f(x)+g(Dx)$
where $x\in \mathbb{R}^n$ and $D\in \mathbb{R}^{m\times n}$. Each
iteration of the linearized ADMM
algorithm~\citep{parikh_proximal_2014} for solving this problem 
has the form
\begin{align}
x & \leftarrow \prox_{\mu f} \left(x - (\mu/\rho)D^T (D x - z + u )\right)\\
z & \leftarrow \prox_{\rho g} \left(z + u\right) \label{eq:linADMM_steps}\\
u & \leftarrow u + D x - z
\end{align}
 where $z,u\in \mathbb{R}^m$ and the proximal operator is defined as
\begin{equation}
 \prox_{\alpha f}(u) = \min_x \,\, \alpha \cdot f(x)+\frac{1}{2} \norm{ x-u}_2^2.
\label{eq:linADMM_prox}
\end{equation}
Clearly,
\eqref{eq:l1tf_var_st} has this form.\attn{What are $\mu$ and $\rho$?}
% The optimization problem 
% form \eqref{eq:linADMM_opt} by defining
% \begin{align}
% & f(x):= \sum_{k} f_k(x_k) := \sum_{k}x_k+y_{k}^2e^{-x_{k}}\\
% & g(z):= \sum_{l} g_l(z_l) := \sum_{l} \lambda_l |z_l|\label{eq:linADMM_fg}\\
% & z = Dx
% \end{align}
% where $y_k$ is the $k^{th}$ entry of the vector whose entries are
% $y_{ijt}$, and $\lambda_l$ is the $l^{th}$ entry of the vector
% $\Lambda^t=(\lambda_t\textbf{e}_{n_t}^t|\lambda_s\textbf{e}_{n_s}^t)$
% (see \autoref{sec:exten}). 
To perform the steps in \eqref{eq:linADMM_steps}, we need to evaluate
$\prox_{\mu f}$ and $\prox_{\rho g}$. Proximal
algorithms are feasible only if these proximal operators can be
evaluated efficiently which, as we show next, is the case for our
problem.

@@

The curve of the estimated SD captures the periodic variations in the
SD of the signal. Just by looking at this curve, it is hard to say if
the SD is decreasing or increasing. Therefore, we compute the average
of the estimated SD for each year. The estimated SD together with this
annual average is shown in the middle panel of
\autoref{fig:bloom_estimatedSD}. As it can be seen, the annual trend
is not smooth. This is because in the optimization problem
\eqref{eq:l1tf_var}, the smoothness of the annual trend is not
encouraged. To remedy this, we add the following long horizon penalty
to \eqref{eq:l1tf_var}: 
\begin{equation}
\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
\label{eq:lh_penalty}
\end{equation}
 where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$. Also,
 $N_{year}$ is the number of years over which we are performing our
 analysis (here $N_{year}=10$). Since we are working on the weekly
 averaged data, each year corresponds to 52 observations. In the
 matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows
 to the matrix $D$. The estimated SDs using this penalty matrix is
 shown in the right panel of \autoref{fig:bloom_estimatedSD}. The
 annual average of the estimated SDs shows a linear trend with a
 positive slope.

@@

>>>>

%Arash
\usepackage{amsmath,amsthm}
\usepackage{amssymb}
\usepackage{mathrsfs}
\DeclareMathOperator*{\argmin}{argmin}
\usepackage{siunitx}
\usepackage{comment}
\usepackage{mathtools}
\mathtoolsset{showonlyrefs}

@@

\attn{Some equations do not have numbering and some have.}

@@

\attn{Arash: should't the title of this section be "...estimating variance of spatio-temporal data" or something similar?}

@@

Inspired by the $\ell_1$-trend filtering algorithm, we propose a
non-parametric model for estimating the variance of a time-series. To
this end, we assume that at each time step $t$, there is a hidden
variable $h_t$ such that conditioned on $h_t$ the observations $y_t$
are independent normal variables with zero mean and variance
$\exp(h_t)$. The negative log-likelihood of the observed data in this
model is $l(y\given h) \propto -\sum_{t=1}^T h_t - y_t^2e^{-h_t}$. Crucially,
we assume that the hidden variables $h_t$ vary smoothly. To impose
this assumption, we estimate $h_t$ by solving the penalized, negative
log-likelihood: 
\begin{equation}
\min_h -l(y\given h)+\lambda \norm{ Dh }_1
\label{eq:l1tf_var}
\end{equation}
 where $D$ has the same structure as above.

@@

As with~\eqref{eq:l1tf}, one can solve~\eqref{eq:l1tf_var} using the
PDIP algorithm (as in, e.g.,
\texttt{cvxopt}~\citep{andersen_cvxopt:_2013}). First, we note that
this is a generalized LASSO problem
\citep{tibshirani_solution_2011}. The dual of a generalized LASSO with
the objective $f(x)+\lambda \norm{ Dx }_1$ is:

\begin{align}
\min_\nu&\quad f^*(-D^\top\nu) & \mbox{s.t.}&\quad \norm{ \nu }_\infty \le \lambda
\end{align}
 where $f^*(\cdot)$ is the Fenchel conjugate of $f$: $f^*(u)=\max_x
 u^\top x-f(x)$. It is simple to show that 
\begin{equation}
f^*(u)=\sum_t (u_t-1)\log\frac{y_t^2}{1-u_t} + u_t-1.
\label{eq:conj}
\end{equation}
Writing
\begin{equation}
  r_w(v,\mu_1,\mu_2):=
  \begin{bmatrix}
    \nabla f^*(-D^\top v) + D(v-\lambda \one)^\top \mu_1 -
    D(v+\lambda \one)^\top \mu_2\\
    -\mu_1(v-\lambda\one)+\mu_2(v + \lambda\one) -w^{-1}\one
  \end{bmatrix}
  =
  \begin{bmatrix}
  0\\0\end{bmatrix}
\label{eq:resid}
\end{equation}

for $w>0$, where $\mu_1$ and $\mu_2$ are dual variables for the
$\ell_\infty$ constraint, as $w\rightarrow\infty$,
the solution to this nonlinear system reduces to the KKT
conditions. Therefore, the PDIP method takes Newton steps to solve the
system for a series of increasing values of $w$. \autoref{alg:pdip}
provides the details. 
\begin{algorithm}[tb]
  \caption{PDIP for $\ell_1$ variance estimation}
  \label{alg:pdip}
  \begin{algorithmic}
    \REQUIRE $\lambda>0$, $w>0$, $\nu\leftarrow 0$, $\mu_1\leftarrow
    0$, $\mu_2\leftarrow 0$, $J\in\mathbb{Z}^+$, $\{w_k\}$
    \COMMENT{Initialization} 
    \FOR[Central path]{$k=1,2,\ldots$}
    \FOR[Newton updates]{$j=1,\ldots,J$}
    \STATE Solve $A [\Delta\nu\; \Delta\mu_1\; \Delta\mu_2]^\top=r_{w_k}$ to find the search direction.
    \STATE $\quad\quad A$ is the Jacobian of $r_w$ in \eqref{eq:resid}.
    \STATE Update $[\nu^{j+1}\; \mu^{j+1}_1\; \mu^{j+1}_2] \leftarrow
    [\nu^{j}\; \mu^{j}_1\; \mu^{j}_2] + [\Delta\nu\; \Delta\mu_1\;
    \Delta\mu_2]$. 
    \ENDFOR
   \ENDFOR
   \RETURN $h=\log\frac{y^2}{1+D^\top\nu}$ \attn{Is this right?
     Explicit form of $A$}
  \end{algorithmic}
\end{algorithm}

@@

The first term in the objective is proportional to the negative
log-likelihood, the second is the temporal penalty for the
time-series at each location $(i,j)$, while the third and fourth,
penalize the difference between the estimated variance of two
vertically and horizontally adjacent points, respectively. The spatial
component of this
penalty is a special case of trend
filtering on graphs~\citep{wang_trend_2016} which penalizes the difference
between the estimated values of the signal on the connected
nodes. As before, we can write
\eqref{eq:l1tf_var_st} in matrix form where $h$ is an $T\times
S$ vector and $D$ is replaced by $D_{TS} \in \mathbb{R}^{(N_t+N_s) \times (T \cdot S)	}$, where $N_t=S \cdot (T-2)$ and $N_s=T \cdot (2n_rn_c-n_r)$ are the number of temporal and spatial constraints, respectively\footnote{$N_s$ is obtained by counting the number of unique constraints at each location and at all times.}. Then, as we
have two different tuning parameters for the temporal and spatial
components, we write $\Lambda =\left[\lambda_1\one_{N_t}^\top,\;
  \lambda_2\one_{N_s}^\top\right]^\top$ leading to \footnote{Throughout the paper, we use $|x|$ for both scalars and vectors. For vectors we use this to denote a vector obtained by taking the absolute value of each entry of $x$.}:
% let $h$ be a
% vector whose first $T$ entries are $h_{11t}$ for $t=1,...,T$, its next
% $T$ entries are $h_{21t}$ and so on, then the optimization problem in
% the matrix form is as
\begin{equation}
\min_h -y\given h)+ \Lambda^\top | D_{TS}h |
\label{eq:l1tf_var_st_mat}
\end{equation}
% where $\Lambda = \mathbf{I}_{Tn_rn_c}\left[\lambda_1\one_T^\top,\;
%   \lambda_2\one_{n_rn_c}^\top\right]$ and 
% $D_{total}^t=[D_{t}^t|D_{s}^t]$, where $n_t$ and $n_s$ are the number
% of rows of $D_{t}$ and $D_{s}$, respectively (see below). The matrix
% $D_t$ is the following block-diagonal matrix and corresponds to the
% temporal penalty: 
% \begin{equation}
% D_t=\begin{bmatrix}
% D &  & \\ 
%  & \ddots & \\
%  &  & D
% \end{bmatrix}
% \label{eq:d_t_matrix}
% \end{equation}
%  where $D$ was defined in \eqref{eq:d_matrix}. The number of the
%  diagonal blocks is equal to the grid size $n_r \times n_c$. Each row
%  of the matrix $D_s$ corresponds to one spatial constraint in
%  \eqref{eq:l1tf_var_st}. For example, the first $T$ rows correspond to
%  $|h_{11t}-h_{21t}|$ for $t=1,...,T$, the next $T$ rows correspond to
%  $|h_{11t}-h_{12t}|$, and so on. 
% \attn{Detailed discussion of the PDIP to here}

@@

For a spatial grid of size $S$ and $T$ time steps, $D_{ST}$ will have
$3Tn_rn_c-2n_rn_c-Tn_r$ rows and $ST$ columns. For a $1^\circ\times
1^\circ$ grid over the
entire northern hemisphere and daily data over 10 years, we have
$n_r=90$, $n_c=180$, $T=3650$ and so $D_{ST}$ has approximately
$10^8$ columns and $10^8$
rows. In each
step of \autoref{alg:pdip}, we need to solve a linear system of
equations in $A$ which depends on $D_{ST}^\top D_{ST}$ (see \citep{boyd_convex_2004}
equation 11.54). Therefore, 
applying the PDIP directly is infeasible for our data.\footnote{We note that this
  is a highly structured and sparse matrix, but, unlike trend
  filtering alone, it is not banded. We are unaware of general
  linear algebra techniques for inverting such matrix, despite our
  best efforts.}

@@

\begin{comment}
General (unconstrained) optimization problems $\min_z
\sum_i f_i(z)$, $z\in\mathbb{R}^n$ can be rewritten equivalently using a
collection of local variables, $x_i\in \mathbb{R}^{n_i}$, and a
constraint. Following the notation of 
\citep{boyd_distributed_2011}, let $k=\mathscr{G}(i,j)$ which
means that the $j^\text{th}$ entry of $x_i$ is $z_k$ (or
$(x_i)_j=z_k$) and define $\tilde{z}_i \in \mathbb{R}^{n_i}$ by
$(\tilde{z}_i)_j=(x_i)_j$. Then the original unconstrained
optimization problem is equivalent to the following constrained
problem
$\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)$ subject to $\tilde{z}_i=x_i$.
Now, we can apply ADMM to the augmented Lagrangian of this
problem: 
\begin{align}
x_i&\leftarrow \argmin_{x_i} f_i(x_i) + (u_i)^\top x_i +
     (\rho/2) \norm{ x_i-\tilde{z}_i }_2^2\\
%\label{eq:ADMM_steps}
z_k&\leftarrow (1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i)_j\\
u_i&\leftarrow u_i + \rho (x_i-\tilde{z}_i)
\end{align}
\end{comment}

Consider an optimization problem of the form $\min_h f(h)$, where $h\in\mathbb{R}^n$ is the \textit{global variable} and $f(h):\mathbb{R}^n \rightarrow \mathbb{R}\cup \{+\infty\}$ is convex. The goal in consensus optimization is to break this problem into several smaller sub-problems which can be solved independently in each iteration of optimization.

Assume that it is possible to define a set of \textit{local variables} $x_i \in \mathbb{R}^{n_i}$ such that $f(h)=\sum_i f_i(x_i)$, where each $x_i$ is a subset of the global variable $h$. More specifically, each entry of the local variables corresponds to an entry of the global variable. Therefore we can define a mapping $\mathscr{G}(i,j)$ from the local variables indices into the global variable indices: $k=\mathscr{G}(i,j)$ means that the $j^\text{th}$ entry of $x_i$ is $h_k$ (or $(x_i)_j=h_k$). For ease of notation, define $\tilde{h}_i \in \mathbb{R}^{n_i}$ as $(\tilde{h}_i)_j=h_{\mathscr{G}(i,j)}$. Then, the original optimization problem is equivalent to the following problem:

\begin{equation}
\begin{aligned}
\min_{\{x_1,...,x_N  \}} \sum_i f_i(x_i)\\
 s.t. \,\,\, \tilde{h}_i=x_i \,\,\,\,\,\,
\end{aligned}
\label{eq:consADMM}
\end{equation}

It is important to note that each entry of the global variable may correspond to several entries of the local variables and so the constraints $\tilde{h}_i=x_i$ enforce the consensus between the local variables corresponding to the same global variable.

The \textit{augmented Lagrangian} corresponding to the problem \ref{eq:consADMM} is $L_\rho(x,h,y)=\sum_i \big(f_i(x_i)+u_i^\top(x_i-\tilde{h}_i) + (\rho/2) \lVert x_i-\tilde{h}_i \lVert_2^2 \big)$. Now, we can apply ADMM to $L_\rho$ which results in the following ADMM updating steps at each iteration $m$:

\begin{equation}
\begin{aligned}
& x_i^{m+1}:=\argmin_{x_i} \bigg( f_i(x_i) + (u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{h}_i^m \lVert_2^2\bigg)\\
& h_k^{m+1}:=(1/S_k)\sum_{\mathscr{G}(i,j)=k} (x_i^{m+1})_j\\
& u_i^{m+1}:=u_i^m + \rho (x_i^{m+1}-\tilde{h}_i^{m+1})
\end{aligned}
\label{eq:ADMM_steps}
\end{equation}

Here, $S_k$ is the number of local variable entries that correspond to
$h_k$, and $u_i$ are the Lagrange multipliers.

\begin{comment}
The $x$-update is
often a simple modification as the original optimization, though over only a small
partition, the $h$-update simply averages over collections of
analogous local variables while the $u$ update enforces sharing. Both
the $x$ and $u$ updates are parallelizable over the partition of local
variables.
\end{comment}

\attn{I don't much like this explanation or notation. Seems unclear}

\attn{Arash: I re-wrote this section.}

To solve the optimization problem \eqref{eq:l1tf_var_st_mat} or
\eqref{eq:l1tf_var_st} using this method, we need to address two
questions: first, how to choose the 
local variables $x_i$, and second, how to solve the optimization
problem for updating these variables (the first line of
\eqref{eq:ADMM_steps} which we will refer to it as \textit{x-update step}).

In \autoref{fig:data_cube}, the global variable $h$ is represented
as a cube (using the subset of the US as an example). We decompose $h$
into sub-cubes as shown in the figure by white lines. It is easy to see
that by this definition of $x_i$, the objective
\eqref{eq:l1tf_var_st_mat} decomposes as $\sum_i f_i(x_i)$ where
$f_i(x_i)=-l(y_i\given x_i)+\Lambda_{(i)}^\top |D_{(i)}x_i|$, and
$\Lambda_{(i)}$ and $D_{(i)}$ contain the temporal and spatial
penalties corresponding to $x_i$ only. The x-update step in \autoref{eq:ADMM_steps} is the following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) +
(u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$.

\attn{I would like to be able to write all this in terms of $h$ instead of $x$}

\attn{Arash: now I'm using $h$ as the global variables. We cannot write the local variables in terms of $h$ since they are different from $h$.}

\begin{figure}[tb]
  \centering
  \includegraphics[height=.2\textheight]{Figures/data_cube}
    \caption{The cube represents the global variable $h$ in space and
      time. The sub-cubes specified by the white lines are
      $x_i$. \attn{Can we use $h_i$ here instead of $x_i$?}}
    \label{fig:data_cube}
\end{figure}

By this definition of $x_i$, the update step for $x_i$ is the
following optimization problem: $x_i^{m+1}:=\argmin_{x_i} ( f_i(x_i) +
(u_i^m)^\top x_i + (\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2)$. We solve this using the PDIP method. As it was explained in \autoref{sec:l1tf_var}, to use PDIP we need to compute the dual problem, which in turn needs the computation of the Fenchel conjugate of the loss function. In addition, referring to \autoref{alg:pdip}, we need to compute the gradient and Jacobian of the conjugate functions. It is important to note that compare to the optimization problem \autoref{eq:l1tf_var}, the loss function in the x-update step in \autoref{eq:ADMM_steps} includes the quadratic term $(\rho/2) \lVert x_i-\tilde{z}_i^m \lVert_2^2$. This makes the computation more involved than \autoref{sec:l1tf_var}. The details of the computations are explained in Appendix 1.

\attn{I'm not sure if we want to devote some space to write this as a separate algorithm. The steps are those mentioned in \ref{eq:ADMM_steps} and the x-update is now explained in the appendix. So we may want to remove algorithm \ref{alg:ADMM}.}

@@

\attn{Need a good dummy variable here, it can't be $u$.} Consider the generic optimization problem
$\min_x f(x)+g(Dx)$
where $x\in \mathbb{R}^n$ and $D\in \mathbb{R}^{m\times n}$. Each
iteration of the linearized ADMM
algorithm~\citep{parikh_proximal_2014} for solving this problem 
has the form
\begin{align}
x & \leftarrow \prox_{\mu f} \left(x - (\mu/\rho)D^\top (D x - z + u )\right)\\
z & \leftarrow \prox_{\rho g} \left(z + u\right) \label{eq:linADMM_steps}\\
u & \leftarrow u + D x - z
\end{align}
 where $z,u\in \mathbb{R}^m$ and the proximal operator is defined as
\begin{equation}
 \prox_{\alpha f}(u) = \min_x \,\, \alpha \cdot f(x)+\frac{1}{2} \norm{ x-u}_2^2.
\label{eq:linADMM_prox}
\end{equation}
Clearly,
\eqref{eq:l1tf_var_st} has this form.\attn{What are $\mu$ and $\rho$?}
% The optimization problem 
% form \eqref{eq:linADMM_opt} by defining
% \begin{align}
% & f(x):= \sum_{k} f_k(x_k) := \sum_{k}x_k+y_{k}^2e^{-x_{k}}\\
% & g(z):= \sum_{l} g_l(z_l) := \sum_{l} \lambda_l |z_l|\label{eq:linADMM_fg}\\
% & z = Dx
% \end{align}
% where $y_k$ is the $k^{th}$ entry of the vector whose entries are
% $y_{ijt}$, and $\lambda_l$ is the $l^{th}$ entry of the vector
% $\Lambda^\top=(\lambda_t\textbf{e}_{n_t}^\top|\lambda_s\textbf{e}_{n_s}^\top)$
% (see \autoref{sec:exten}). 
To perform the steps in \eqref{eq:linADMM_steps}, we need to evaluate
$\prox_{\mu f}$ and $\prox_{\rho g}$. Proximal
algorithms are feasible only if these proximal operators can be
evaluated efficiently which, as we show next, is the case for our
problem.

@@

The curve of the estimated SD captures the periodic variations in the
SD of the signal. Just by looking at this curve, it is hard to say if
the SD is decreasing or increasing. Therefore, we compute the average
of the estimated SD for each year. The estimated SD together with this
annual average is shown in the middle panel of
\autoref{fig:bloom_estimatedSD}. As it can be seen, the annual trend
is not smooth. This is because in the optimization problem
\eqref{eq:l1tf_var}, the smoothness of the annual trend is not
encouraged. To remedy this, we add the following long horizon penalty
to \eqref{eq:l1tf_var}:

\begin{equation}
\sum_{i=1}^{N_{year}-2} \Big\arrowvert \sum_{t=1}^{52} h_{t_1}-2h_{t_2}+h_{t_3}  \Big\arrowvert
\label{eq:lh_penalty}
\end{equation}

where $t_1=52(i-1)+t$, $t_2=52i+t$ and $t_3=52(i+1)+t$. Also, $N_{year}$ is the number of years over which we are performing our analysis (here $N_{year}=10$). Since we are working on the weekly averaged data, each year corresponds to 52 observations. In the matrix form, the penalty \eqref{eq:lh_penalty} adds $N_{year}$ rows to the matrix $D$. The estimated SDs using this penalty matrix is shown in the right panel of \autoref{fig:bloom_estimatedSD}. The annual average of the estimated SDs shows a linear trend with a positive slope.

@@

\section{Appendix}

\attn{put this in nips appendix format}

In this Appendix we give more details on performing the x-update step in \autoref{eq:ADMM_steps}. We need to solve the following optimization problem:

\begin{equation}
\hat{x}:=\argmin_{x} \bigg( \sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2 + \Lambda^\top |D x| \bigg)
\label{eq:x_update_opt}
\end{equation}

\noindent where $n_b$ is the number of local variables in each sub-cube in \autoref{fig:data_cube}, and for ease of notation we have dropped the subscript $i$ and superscript $m$. Let $f(x)=\sum_{j=1}^{n_b} (x_j + y_j^2e^{-x_j}) + (\rho/2) \lVert x-\tilde{z} + u \lVert_2^2$. As it was explained in \autoref{sec:l1tf_var}, the dual of this optimization problem is: $\min_\nu f^*(-D^\top\nu)$ with the constraints $|\nu_k| \le \Lambda_k$. So to use PDIP we first need to compute the conjugate function $f^*(\cdot)$. We have:
